Slavnov products and KP tau functions

### 2025-05-02 draft.tex
finish paper

\paragraph{\emph{Proof}} Let us first use that
\begin{equation}
\label{eq:scalar-prod}
  \bra{\Psi(\bm{v})} \left(\mathcal{T}(u_j)\ket{\Psi(\bm{u}_j)} \right) =
  \left(\bra{\Psi(\bm{v})}\mathcal{T}(u_j)\right)\ket{\Psi(\bm{u}_j)}
  \qquad j =1, \dots, n+1\; .
\end{equation}
The right-hand side of this equation can be simplified with the eigenvalue
expression~(\ref{eq:eigenvalue-bethe}), that is
\begin{equation}
  \left(\bra{\Psi(\bm{v})}\mathcal{T}(u_j)\right)\ket{\Psi(\bm{u}_j)}  =
\Lambda(u_j; \bm{v}) \bracket{\Psi(\bm{v})}{\Psi(\bm{u}_j)} = 
\Lambda(u_j; \bm{v}) \bm{\zeta}_j\; . 
\end{equation}

@@

This section deals with the properties of the Baker-Akhiezer functions
associated to the above tau functions. There are some interesting
aspects of these tau functions that deserve further analysis, and here
we just touch some of these aspects.

Let us first start with some rewriting of the tau functions above.

@@

We now want to write the Slavnov products we defined above in terms of
the following coordinates
\begin{equation}
  t_p = \frac{1}{p}\sum_{j=1}^n z_j^p\; ,
\end{equation}
the so called, Miwa coordinates.

Let us also define
\begin{equation}
  \begin{split}
    e^{\xi(\bm{t}, \lambda)} & = \exp\left( \sum_{p=1}^\infty t_p \lambda^p \right)
    = \exp\left( \sum_{p=1}^\infty \sum_{j=1}^n \frac{1}{p} z^p \lambda^p \right) \\
    & = \exp\left( \sum_{j=1}^n \sum_{p=1}^\infty  \frac{1}{p} z_j^p \lambda^p \right)
    = \exp\left( - \sum_{j=1}^n \ln( 1 -  z_j \lambda ) \right) \\
    & =  \prod_{j=1}^n \frac{1}{1 -  z_j \lambda} 
  \end{split}
\end{equation}
where \(\lambda\) is a complex parameter. 
and we see that these functions have simple poles at \( \lambda = z_j^{-1}\).

\begin{proposition}
The tau function~(\ref{eq:tau-function}) admits an integral
representation, see ~\cite{Araujo:2021ghu} and references therein,
of the form
\begin{subequations}
\begin{equation}
  \label{eq:tau-function-miwa}
  \bm{\tau}(\bm{t}, \bm{v}) =
    \det_{j,k}\left(
    \oint_{\gamma_k} \frac{d w}{2\pi i} e^{\xi(\bm{t}, w^{-1})} \frac{w^{-j}\mathcal{Y}_k(w)}{w - v_k} \right)\; ,
\end{equation}
where 
\begin{equation}
  \xi(\bm{t}, w^{-1}) = \sum_{p=1}^\infty t_p w^{-p} \quad \Leftrightarrow \quad
  \xi(\bm{z}, w^{-1}) = w^n \prod_{j=1}^n (w - z_j)^{-1} \; . 
\end{equation}
In the above integral, we consider that integration curve \(\gamma_k\)
encloses all poles except the Bethe root \(v_k\).
\end{subequations}

In terms of the \(z\)-coordinates, the tau functions becomes
\begin{equation}
  \label{eq:tau-function}
  \bm{\tau}(\bm{z}, \bm{v}) =
    \det_{j,k}\left(
    \oint_{\gamma_k} \frac{d w}{2\pi i}
    \prod_{l=1}^n \frac{1}{w - z_l} \frac{w^{n-j}\mathcal{Y}_k(w)}{w - v_k} \right) \; . 
\end{equation}
Now, it is easy to see that \(w =0\) is not a pole.

All in all, the integration curve \(\gamma_k\) encloses all poles
\(\{z_j\}_{j=1}^n\).  One of the advantages of this representation is
that now it makes sense to consider the limit of the infinite chain,
\(n\to \infty\).
\end{proposition}

\paragraph{\emph{Proof}}
There are some minor differences between our calculations and the
original references; namely, the integration contour \(\gamma_k\) and
sign conventions. For this reason, we want to show
explicitly this result.

@@

It is easy to see that
\begin{equation}
  \begin{split}
  \det \bm{\mathcal{K}} & = \det(\bm{\Omega}) \det_{l,j}\left( \frac{z_l^{n-j}}{\prod_{j\neq l} (z_l - z_j)}\right) \\
  & = \det(\bm{\Omega}) \det_{l, j}(z_l^{n-j}) \prod_{\substack{j, l =1 \\ j \neq l}}^n (z_l - z_j)^{-1} \; .
  \end{split}
\end{equation}
We also have the identities 
\begin{subequations}
  \begin{equation}
	\det_{l,j}(z_l^{n-j}) = (-1)^{n(n-1)/2} \Delta(\bm{z})\; ,
  \end{equation}
  and
  \begin{equation}
	\prod_{\substack{j, l = 1 \\ j \neq l}}^n (z_l - z_j) = (-1)^{n(n-1)/2} \Delta(\bm{z})^2\; . 
  \end{equation}
\end{subequations}
Putting all these facts together, we finally see that 
\(\bm{\tau}(\bm{z}, \bm{v}) = \det\bm{\mathcal{K}}\). \qed

@@

The Baker-Akhiezer funcion is defined in terms of the Japanese formula~\cite{Zabrodin2018}
\begin{equation}
  \psi(\bm{t},\bm{v}; \lambda) = e^{\xi(\bm{t}, \lambda)}
  \frac{\bm{\tau}(\bm{t} - [\lambda^{-1}], \bm{v})}{\bm{\tau}(\bm{t}, \bm{v})}
\end{equation}
where 
\begin{equation}
  \bm{t} - [\lambda^{-1}] = 
  \{t_1 - \lambda^{-1}, t_2 - \lambda^{-2}/ 2 ,  t_2 - \lambda^{-3}/3, \dots,  t_p - \lambda^{-p}/p, \dots \}\; .
\end{equation}

Using the expression~(\ref{eq:tau-function-miwa}), we can write the
Baker-Akhiezer function in terms of the \(z\)-coordinates.
Moreover, 
\begin{equation}
  t_p - \lambda^{-p}/p = \frac{1}{p} \sum_{j=1}^n z_j^ p - \lambda^{-p}/p \; , 
\end{equation}
and we can write
\begin{equation}
  \begin{split}
    e^{\xi(\bm{t} - [\lambda^{-1}], w^{-1})} & = \exp\left[ \sum_{p=1}^\infty \left(t_p  - \frac{\lambda^{-p}}{p}\right) w^{-p} \right]\\
    & = e^{\xi(\bm{t}, w^{-1})} \exp\left[ - \sum_{p=1}^\infty \frac{\lambda^{-p}}{p} w^{-p} \right] \\
    & = e^{\xi(\bm{t}, w^{-1})} \exp\left[ \ln ( 1 - 1 / (\lambda w)) \right]
      = \left( 1 - \frac{1}{\lambda w}\right)  \prod_{j=1}^n \frac{1}{1 -  z_j /w} \\
    & = \frac{w^{n-1}}{\lambda } \left( \lambda w - 1 \right)  \prod_{j=1}^n \frac{1}{w -  z_j} \\
  \end{split}
\end{equation}

Using these expressions, we can write 
\begin{equation}
  \bm{\tau}(\bm{t} - [\lambda^{-1}], \bm{v}) =
  \det_{j,k}\left[
  \oint_{\gamma_k} \frac{dw}{2\pi i} e^{\xi(\bm{t}, w^{-1})} \frac{w^{- j -1} ( \lambda w - 1) \mathcal{Y}_k(w)}{\lambda(w - v_k)} 
  \right]\; .
\end{equation}
that in terms of the \(z\)-coordinates becomes
\begin{equation}
\begin{split}
  \bm{\tau}(\bm{t} - [\lambda^{-1}], \bm{v})
  & = \det_{j,k}\left[\oint_{\gamma_k} \frac{dw}{2\pi i} \prod_{s=1}^n \frac{1}{w - z_s}  
    \frac{( \lambda w - 1) w^{n-j-1}\mathcal{Y}_k(w)}{\lambda(w - v_k)} \right]\\
  & = \det_{j,k}\left[
    \oint_{\gamma_k} \frac{dw}{2\pi i} \prod_{s=1}^n \frac{1}{w - z_s}  
    \frac{w^{n-j-1}\widehat{\mathcal{Y}}_k(w; \lambda)}{(w - v_k)}
    \right]
\end{split}
\end{equation}
where we have defined the function 
\begin{equation}
  \widehat{\mathcal{Y}}_k(w; \lambda)
  = \frac{( \lambda w - 1)}{\lambda} \mathcal{Y}_k(w) \; ,
\end{equation}
From these expressions, it is easy to see that for \(j=n\), then \(w =
0\) is a pole and we must pick this term in the integration curve
\(\gamma_k\).

Let us define the components
\begin{equation}
  \begin{split}
    \widetilde{\mathcal{K}}_{kj} = &
    \oint_{\tilde \gamma_k} \frac{dw}{2\pi i} \prod_{s=1}^n \frac{1}{w - z_s}
    \frac{w^{n-j-1}\widehat{\mathcal{Y}}_k(w; \lambda)}{(w - v_k)}
    + \delta_{nj} F_k(\lambda) \\
    & = \sum_{\ell = 1}^n  \widehat{\Omega}_{k \ell} \left( \frac{z_\ell^{n-j}}{\prod_{s\neq \ell} (z_\ell - z_s)} \right)
    + F_k(\lambda)\delta_{nj}  \\
  \end{split}
\end{equation}
where \(\tilde{\gamma}_k\) is a contour that does not pick
the point \(w=0\); and
\begin{equation}
  \widehat{\Omega}_{k\ell}
  = \frac{1}{z_\ell} \frac{\widetilde{\mathcal{Y}}_k(\lambda, z_\ell)}{z_\ell - v_k}
  = \frac{(\lambda z_\ell -1)}{\lambda z_\ell} {\Omega}_{k\ell} 
\; , \qquad 
 F_k(\lambda) = (-1)^n \frac{\widehat{\mathcal{Y}}_k(\lambda; 0)}{v_k }  \prod_{s=1}^n z_s^{-1} \; .
\end{equation}

@@

Finally, using the elementary symmetric polynomials \(\sigma_p^{(n)}\), we write
\begin{equation}
  \prod_{l=1}^n \left( 1 - \frac{1}{\lambda z_l} \right) = 1 + 
  \sum_{p=1}^n (-1)^p \lambda^{-p} \sigma_p^{(n)}(\bm{z}^{-1})\; .
\end{equation}
where \( \bm{z}^{-1} = \{z_1^{-1}, z_2^{-1}, \dots , z_n^{-1}\}\).

@@

We now want to consider the above calculations in the Miwa
coordinates. This allows us to consider the infinite chain limit.

>>>>

\paragraph{\textbf{\emph{Proof}}} Let us first use that
\begin{equation}
\label{eq:scalar-prod}
  \bra{\Psi(\bm{v})} \left(\mathcal{T}(u_j)\ket{\Psi(\bm{u}_j)} \right) =
  \left(\bra{\Psi(\bm{v})}\mathcal{T}(u_j)\right)\ket{\Psi(\bm{u}_j)}
  \qquad j =1, \dots, n+1\; .
\end{equation}
The right-hand side of this equation can be simplified with the eigenvalue
expression~(\ref{eq:eigenvalue-bethe}), that is
\begin{equation}
  \left(\bra{\Psi(\bm{v})}\mathcal{T}(u_j)\right)\ket{\Psi(\bm{u}_j)}  =
\Lambda(u_j; \bm{v}) \bracket{\Psi(\bm{v})}{\Psi(\bm{u}_j)} = 
\Lambda(u_j; \bm{v}) \bm{\zeta}_j\; . 
\end{equation}

@@

This section deals with some properties of the Baker-Akhiezer
functions associated to the tau functions we found above.  Many
aspects of these functions deserve a thorough examination, but here we
investigagte just the most immediate aspects.

@@

We now want to write the Slavnov products we defined above in terms of
the following coordinates
\begin{equation}
  t_p = \frac{1}{p}\sum_{j=1}^n z_j^p\; ,
\end{equation}
the so called Miwa coordinates.

We define
\begin{equation}
  \xi(\bm{t}, \lambda) = \sum_{p=1}^\infty t_p \lambda^p\; , 
\end{equation}
where \(\lambda\) is a complex parameter. Therefore
\begin{equation}
  \begin{split}
    e^{\xi(\bm{t}, \lambda)} & = \exp\left( \sum_{p=1}^\infty t_p \lambda^p \right)
    = \exp\left( \sum_{p=1}^\infty \sum_{j=1}^n \frac{1}{p} z^p \lambda^p \right) \\
    & = \exp\left( \sum_{j=1}^n \sum_{p=1}^\infty  \frac{1}{p} z_j^p \lambda^p \right)
    = \exp\left( - \sum_{j=1}^n \ln( 1 -  z_j \lambda ) \right) \\
    & =  \prod_{j=1}^n \frac{1}{1 -  z_j \lambda} \; .
  \end{split}
\end{equation}
It is easy to see that these functions have simple poles at \( \lambda
= z_j^{-1}\).

Let us now establish an important result,
see ~\cite{Araujo:2021ghu} and references therein. 
\begin{proposition}
The tau function~(\ref{eq:slav-tau-function}) admits the following integral
representation 
\begin{subequations}
\begin{equation}
  \label{eq:tau-function-miwa}
  \bm{\tau}(\bm{t}, \bm{v}) =
    \det_{j,k}\left(
    \oint_{\gamma_k} \frac{d w}{2\pi i} e^{\xi(\bm{t}, w^{-1})} \frac{w^{-j}\mathcal{Y}_k(w)}{w - v_k} \right)\; ,
\end{equation}
where 
\begin{equation}
  \xi(\bm{t}, w^{-1}) = \sum_{p=1}^\infty t_p w^{-p} \quad \Leftrightarrow \quad
  \xi(\bm{z}, w^{-1}) = w^n \prod_{j=1}^n (w - z_j)^{-1} \; . 
\end{equation}
In the above integral, we consider that integration curve \(\gamma_k\)
encloses all poles except the Bethe root \(v_k\).
\end{subequations}
\end{proposition}

\paragraph{\textbf{\emph{Proof}}}
In terms of the \(z\)-coordinates, the tau functions becomes
\begin{equation}
  \label{eq:tau-function}
  \bm{\tau}(\bm{z}, \bm{v}) =
    \det_{j,k}\left(
    \oint_{\gamma_k} \frac{d w}{2\pi i}
    \prod_{l=1}^n \frac{1}{w - z_l} \frac{w^{n-j}\mathcal{Y}_k(w)}{w - v_k} \right) \; . 
\end{equation}
Now, it is easy to see that \(w =0\) is not a pole.  All in all, the
integration curve \(\gamma_k\) encloses points \(\{z_j\}_{j=1}^n\).

@@

It is easy to see that
\begin{equation}
  \begin{split}
  \det \bm{\mathcal{K}} & = \det(\bm{\Omega}) \det_{l,j}\left( \frac{z_l^{n-j}}{\prod_{j\neq l} (z_l - z_j)}\right) \\
  & = \det(\bm{\Omega}) \det_{l, j}(z_l^{n-j}) \prod_{\substack{j, l =1 \\ j \neq l}}^n (z_l - z_j)^{-1} \; .
  \end{split}
\end{equation}
We also have the identities 
\begin{subequations}
  \begin{equation}
	\det_{l,j}(z_l^{n-j}) = (-1)^{n(n-1)/2} \Delta(\bm{z})\; ,
  \end{equation}
  and
  \begin{equation}
	\prod_{\substack{j, l = 1 \\ j \neq l}}^n (z_l - z_j) = (-1)^{n(n-1)/2} \Delta(\bm{z})^2\; . 
  \end{equation}
\end{subequations}
Putting all these facts together, we finally see that 
\(\bm{\tau}(\bm{z}, \bm{v}) = \det\bm{\mathcal{K}}\), that
is preciselly the expression~(\ref{eq:slav-tau-function}).\\
\qed

One of the advantages of this representation is
that now it is easier to consider the limit of the infinite chain,
\(n\to \infty\).

@@

The Baker-Akhiezer (BA) funcion is defined in terms of the Japanese formula~\cite{Zabrodin2018}
\begin{equation}
  \psi(\bm{t},\bm{v}; \lambda) = e^{\xi(\bm{t}, \lambda)}
  \frac{\bm{\tau}(\bm{t} - [\lambda^{-1}], \bm{v})}{\bm{\tau}(\bm{t}, \bm{v})}
\end{equation}
where 
\begin{equation}
  \bm{t} - [\lambda^{-1}] = 
  \{t_1 - \lambda^{-1}, t_2 - \lambda^{-2}/ 2 ,  t_2 - \lambda^{-3}/3, \dots,  t_p - \lambda^{-p}/p, \dots \}\; .
\end{equation}

\subsubsection{BA in \(z\)-coordinates}
Using the expression~(\ref{eq:tau-function-miwa}), we can write the
Baker-Akhiezer function in terms of the \(z\)-coordinates.
Let us first write
\begin{equation}
  t_p - \lambda^{-p}/p = \frac{1}{p} \sum_{j=1}^n z_j^ p - \lambda^{-p}/p \; , 
\end{equation}
therefore
\begin{equation}
  \begin{split}
    e^{\xi(\bm{t} - [\lambda^{-1}], w^{-1})} & = \exp\left[ \sum_{p=1}^\infty \left(t_p  - \frac{\lambda^{-p}}{p}\right) w^{-p} \right]\\
    & = e^{\xi(\bm{t}, w^{-1})} \exp\left[ - \sum_{p=1}^\infty \frac{\lambda^{-p}}{p} w^{-p} \right] \\
    & = e^{\xi(\bm{t}, w^{-1})} \exp\left[ \ln ( 1 - 1 / (\lambda w)) \right]
      = \left( 1 - \frac{1}{\lambda w}\right)  \prod_{j=1}^n \frac{1}{1 -  z_j /w} \\
    & = \frac{w^{n-1}}{\lambda } \left( \lambda w - 1 \right)  \prod_{j=1}^n \frac{1}{w -  z_j} \\
  \end{split}
\end{equation}

Using these expressions, we can write 
\begin{equation}
  \bm{\tau}(\bm{t} - [\lambda^{-1}], \bm{v}) =
  \det_{j,k}\left[
  \oint_{\gamma_k} \frac{dw}{2\pi i} e^{\xi(\bm{t}, w^{-1})} \frac{w^{- j -1} ( \lambda w - 1) \mathcal{Y}_k(w)}{\lambda(w - v_k)} 
  \right]\; .
\end{equation}
that in terms of the \(z\)-coordinates becomes
\begin{equation}
\begin{split}
  \bm{\tau}(\bm{t} - [\lambda^{-1}], \bm{v})
  & = \det_{j,k}\left[\oint_{\gamma_k} \frac{dw}{2\pi i} \prod_{s=1}^n \frac{1}{w - z_s}  
    \frac{( \lambda w - 1) w^{n-j-1}\mathcal{Y}_k(w)}{\lambda(w - v_k)} \right]\\
  & = \det_{j,k}\left[
    \oint_{\gamma_k} \frac{dw}{2\pi i} \prod_{s=1}^n \frac{1}{w - z_s}  
    \frac{w^{n-j-1}\widehat{\mathcal{Y}}_k(w; \lambda)}{(w - v_k)}
    \right]
\end{split}
\end{equation}
where we have defined the function 
\begin{equation}
  \widehat{\mathcal{Y}}_k(w; \lambda)
  = \frac{( \lambda w - 1)}{\lambda} \mathcal{Y}_k(w) \; ,
\end{equation}
From these expressions, it is easy to see that for \(j=n\) the point
\(w = 0\) is a pole; and we must consider it in the integration curve \(\gamma_k\).

It is convenient to define the components
\begin{equation}
  \begin{split}
    \widetilde{\mathcal{K}}_{kj} = &
    \oint_{\tilde \gamma_k} \frac{dw}{2\pi i} \prod_{s=1}^n \frac{1}{w - z_s}
    \frac{w^{n-j-1}\widehat{\mathcal{Y}}_k(w; \lambda)}{(w - v_k)}
    + \delta_{nj} F_k(\lambda) \\
    & = \sum_{\ell = 1}^n  \widehat{\Omega}_{k \ell} \left( \frac{z_\ell^{n-j}}{\prod_{s\neq \ell} (z_\ell - z_s)} \right)
    + F_k(\lambda)\delta_{nj}  \\
  \end{split}
\end{equation}
where \(\tilde{\gamma}_k\) is a deformation contour that does not pick
the point \(w=0\); and
\begin{equation}
  \widehat{\Omega}_{k\ell}
  = \frac{1}{z_\ell} \frac{\widetilde{\mathcal{Y}}_k(\lambda, z_\ell)}{z_\ell - v_k}
  = \frac{(\lambda z_\ell -1)}{\lambda z_\ell} {\Omega}_{k\ell} 
\; , \qquad 
 F_k(\lambda) = (-1)^{n+1} \frac{\widehat{\mathcal{Y}}_k(0; \lambda)}{v_k }  \prod_{s=1}^n z_s^{-1} \; .
\end{equation}

@@

Finally, using the elementary symmetric polynomials \(\sigma_p^{(n)}\), we write
\begin{equation}
  \prod_{l=1}^n \left( 1 - \frac{1}{\lambda z_l} \right) = 1 + 
  \sum_{p=1}^n (-1)^p \lambda^{-p} \sigma_p^{(n)}(\bm{z}^{-1})\; , 
\end{equation}
where \( \bm{z}^{-1} = \{z_1^{-1}, z_2^{-1}, \dots , z_n^{-1}\}\).

@@

We now want to consider the above calculations in the Miwa
coordinates.
